---
title: "Building and Processing BiocBuildDB Reports"
author: "Sean Davis"
date: 2025-12-31
format: html
vignette: >
  %\VignetteIndexEntry{Building and Processing BiocBuildDB Reports}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

## Overview

The `BiocBuildDB` package provides functionality to process build reports
generated by the Bioconductor build system. These reports are typically
created during the continuous integration and testing of Bioconductor packages.


::: {.callout-note}
This vignette documents INTERNAL functionality of the `BiocBuildDB` and is,
in general, not intended for end users.

In general, the build report processing is not something that an end user will
need to pay any attention to. However, the automation and processing scripts
are also included in the package directory and detailed below.
:::

## Build report processing

To process all new build reports (those that changed since the last processing),
you can use the following code:

```{r eval=FALSE}
library(BiocBuildDB)
reportdb_filename = 'reportdb.csv'
dir.create('report_dir')
process_all_new_reports(reportdb_filename, 'report_dir')
```


If the `reportdb.csv` file does not exist, it will be created. If it does exist,
it will be read and updated after successfully localizing (copying) the
new report.tgz files to the `report_dir` directory.

The `process_all_new_reports` function will untar each report.tgz file and
create a set of tables from the report directory. The tables will be written
to files in the `report_dir` directory with the md5 hash of the report directory
as a prefix. 

The result will be a `report_dir` directory with a set of csv files containing
the tables from the report directories as well as the `report.tgz` files. 
Related files are named with the same md5 hash prefix of the report.tgz file. 

An example resulting directory might look like:

```
-rw-r--r--  1 seandavis  staff   472K Feb  2 11:20 2cc2a659a35d607f71655f3c9c9e4283-build_summary.csv.gz
-rw-r--r--  1 seandavis  staff    87K Feb  2 11:20 2cc2a659a35d607f71655f3c9c9e4283-info.csv.gz
-rw-r--r--  1 seandavis  staff    37K Feb  2 11:20 2cc2a659a35d607f71655f3c9c9e4283-propagation_status.csv.gz
-rw-r--r--  1 seandavis  staff    50M Feb  2 11:18 2cc2a659a35d607f71655f3c9c9e4283-report.tgz
-rw-r--r--  1 seandavis  staff   2.9K Feb  2 11:20 2e6b75f554d439ba3dc993e77862a973-build_summary.csv.gz
-rw-r--r--  1 seandavis  staff   2.2K Feb  2 11:20 2e6b75f554d439ba3dc993e77862a973-info.csv.gz
-rw-r--r--  1 seandavis  staff   514B Feb  2 11:20 2e6b75f554d439ba3dc993e77862a973-propagation_status.csv.gz
-rw-r--r--  1 seandavis  staff   202K Feb  2 11:18 2e6b75f554d439ba3dc993e77862a973-report.tgz
-rw-r--r--  1 seandavis  staff   497K Feb  2 11:20 354e509ee0e71215f7669fda8bad0246-build_summary.csv.gz
-rw-r--r--  1 seandavis  staff    94K Feb  2 11:20 354e509ee0e71215f7669fda8bad0246-info.csv.gz
-rw-r--r--  1 seandavis  staff    40K Feb  2 11:20 354e509ee0e71215f7669fda8bad0246-propagation_status.csv.gz
-rw-r--r--  1 seandavis  staff    70M Feb  2 11:18 354e509ee0e71215f7669fda8bad0246-report.tgz
```


**After running this, you may want to sync the `report_dir` directory with a cloud
storage service such as Amazon S3 or Google Cloud Storage for safe keeping.**


## Work with a report.tgz file (just FYI)

You shouldn't need to use this functionality directly, but it is here
to show how the package works and some example output in the tables.

Show an example of how to work with a report.tgz file.

```{r}
library(BiocBuildDB)
report_tgz <- example_report_tgz()
report_dir <- untar_report_tgz(report_tgz)
summary_df <- get_build_summary_table(report_dir)
info_df <- get_info_table(report_dir)
prop_df <- get_propagation_status_table(report_dir)
```

Show the first few rows of each table.

```{r}
head(summary_df)
colnames(summary_df)
head(info_df)
colnames(info_df)
head(prop_df)
colnames(prop_df)
```

Present a histogram of build times (in seconds) for the packages
in the example build report.

```{r}
hist(as.numeric(summary_df$endedat - summary_df$startedat))
```

